Variational Monte Carlo on bosonic systems

### 2018-04-01 project_1/article/article.tex
Introduction

>>>>

    By building up an increasingly intricate system we eloquently illustrate the
    strength of the VMC methods. Any Monte Carlo method draws on the fact that
    the computer will not complain when being tasked to perform a very similar
    task up to several million times in rapid succession. In general Monte Carlo
    methods is in essence a series of guesses where the current guess is only a minor 
    alteration of the one prior to it. The trick lies in only accepting the subsequent 
    guesses that improves upon our configuration up to a certain likelihood.

    The method we employed is called a \emph{Variational} Monte Carlo method, 
    because we make an educated guess at a wavefunction which we allow to contain 
    certain \emph{variational parameters}. We shall see that the system converges to a proper
    \footnote{Global, as far as we can tell.} minimum for only a certain set of parameters.

    First, we start by explaining our theoretical quantum system of a bosonic gas trapped
    in a potential well. Here we also outline the principle under which particles in the gas are
    allowed to interact. Moreover, we introduce the concept of \emph{local energy} and
    then quantum \emph{drift force}. Second, we outline in more detailed algebra the
    nature of a non-interaction system. For this system we are able to derive an exact 
    variational energy to which we can compare the results from our simulations. Third,
    we give a description of the particle-particle interaction implemented as if each 
    particle is a hard sphere. Fourth, we provide a thorough description of algorithms and
    statistical methods employed in this study. Here we introduce the Metropolis-Hastings
    algorithm. This algorithm can by a deft trick be expanded with the method of 
    \emph{importance sampling} which make the way in which we pick the next step in
   	the Monte-Carlo cycling much more intelligent. The main point of note with regards to 
   	the statistical analysis is that the data generated by the Monte Carlo cycling is
   	autocorrelated and computation of variance must be done with the \emph{blocking method}.
   	As an automated way of finding the optimal variational parameter, we sketch out the
   	method of \emph{gradient descent}. Fifth, we put the machinere described in the 
   	previous methods to work. We test against the actual values in a non-interacting system
   	before analysing the more interesting perturbed-trap system with interacting particles. 
   	Lastly, we highlight the more interesting aspects of this study in a discussion of the results
   	before ending on some summary remarks.
